Time delay estimation in unresolved lensed quasars

### 2020-12-22 paper/aa-package/paper.tex
intro

\usepackage[]{natbib}

@@

   \author{G. Wuchterl
          \inst{1}
          \and
          C. Ptolemy\inst{2}\fnmsep\thanks{Just to show the usage
          of the elements in the author field}
          }

@@

   \keywords{giant planet formation --
                $\kappa$-mechanism --
                stability of gas spheres
               }

@@

The Hubble parameter $H_0$ quantifies the expansion rate of the Universe and can be measured in different ways. The Planck collaboration \cite{planck} has found $H_0 = 67.4 \pm 0.5$ km s$^{-1}$ Mpc$^{-1}$ from the Cosmic Microwave Background anisotropies spectrum \footnote{Assuming the $\Lambda$CDM cosmological model.}, whereas, cosmic distance ladder observations \cite{cepheid_H0} give $H_0 = 74.03 \pm 1.42$ km s$^{-1}$ Mpc$^{-1}$, resulting in a tension of about $4.4\sigma$.\\ As first pointed out in \cite{Refsdal1964}, gravitational lensing offers an alternative way of determining the Hubble parameter. The light rays coming from a distant source, e.g. a quasar, can be deflected by the gravitational field of an intervening massive object, like a galaxy. If the field is strong enough, multiple images of the same source will be observed and, by tracking the light intensity of each image over time, one can obtain a so-called \emph{light curve}. The light curves associated with different images will exhibit a mutual time-delay, due to the different paths that the photons have travelled. As shown in \cite{Refsdal1964}, this time-delay can be directly related to $H_0$. The H0LiCOW collaboration \cite{H0licow_XIII} found $H_0 = 73.3_{-1.8}^{+1.7}$ km s$^{-1}$ Mpc$^{-1}$ from a sample of six lensed quasars monitored by the COSMOGRAIL \cite{Cosmograil2020} project, enhancing the tension up to $5.3 \sigma$, and thus motivating further studies aimed at improving the precision in the estimate of $H_0$.

@@

\bibliographystyle{./bibtex/aa} % style aa.bst
\bibliography{./bibtex/biblio}

@@

% All long tables will be placed automatically at the end of the document
%
\longtab{
\begin{landscape}
\begin{longtable}{lllrrr}
\caption{\label{kstars} Sample stars with absolute magnitude}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endfirsthead
\caption{continued.}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endhead
\hline
\endfoot
%%
Gl 33    & 6.37 & K2 V & 7.46 & S & 0.043170\\
Gl 66AB  & 6.26 & K2 V & 8.15 & S & 0.260478\\
Gl 68    & 5.87 & K1 V & 7.47 & P & 0.026610\\
         &      &      &      & H & 0.008686\\
Gl 86
\footnote{Source not included in the HRI catalog. See Sect.~5.4.2 for details.}
         & 5.92 & K0 V & 10.91& S & 0.058230\\
\end{longtable}
\end{landscape}
}
%
%-------------------------------------------------------------
%               Appendices have to be placed at the end, after
%                                        \end{thebibliography}
%-------------------------------------------------------------
\end{thebibliography}

>>>>

\usepackage{natbib}
\bibliographystyle{aa}

@@

   \author{Author1
          \inst{1}
          \and
          Author2\inst{2}\fnmsep\thanks{Just to show the usage
          of the elements in the author field}
          }

@@

   \keywords{Gravitational lensing --
                Hubble parameter --
                Quasars -- 
                Galaxies --
                Machine Learning
               }

@@

The Hubble parameter $H_0$ quantifies the current expansion rate of the Universe. The measured value of $H_0$ from different observations have led to a tension. In particular, EU observations [\cite{planck}] have measured $H_0 = 67.4 \pm 0.5$ km s$^{-1}$ Mpc$^{-1}$, whereas, LU observations [\cite{cepheid_H0}] give $H_0 = 74.03 \pm 1.42$ km s$^{-1}$ Mpc$^{-1}$, resulting in a tension of about $4.4\sigma$.
As first pointed out in [\cite{Refsdal1964}], time-delay cosmography offers an alternative way of determining the Hubble parameter: the light rays coming from a distant source, e.g. a quasar, can be deflected by the gravitational field of an intervening massive object, e.g. a galaxy. If the field is strong enough, multiple images of the same source will be observed and, by tracking the light intensity of each image over time, a \emph{light curve} is obtained. Light curves associated with different images will exhibit a mutual time-delay ($\Delta T$), due to the different paths that the photons have travelled. As shown in [\cite{Refsdal1964}], this time-delay is related to the Hubble parameter as $H_0 \propto 1/\Delta T$. The major results obtained via time-delay cosmography come from the H0LiCOW collaboration [\cite{H0licow_XIII}], which has found $H_0 = 73.3_{-1.8}^{+1.7}$ km s$^{-1}$ Mpc$^{-1}$ from a sample of six lensed quasars monitored by the COSMOGRAIL  project [\cite{Cosmograil2020}], enhancing the tension up to $5.3 \sigma$. However, a more recent analysis of 40 strong gravitational lenses, from TDCOSMO+SLACS \cite{tdcosmo}, has found $H_0 = 67.4_{-3.2}^{+4.1}$ km s$^{-1}$ Mpc$^{-1}$, relaxing the Hubble tension and demonstrating the importance of understanding the lenses mass density profiles. This scenario motivates further studies aimed at improving the precision in the $H_0$ estimation.

@@

%\bibliographystyle{./aa} % style aa.bst
%\bibliographystyle{unsrtnat}
\bibliography{./biblio}

@@

% All long tables will be placed automatically at the end of the document
%
\longtab{
\begin{landscape}
\begin{longtable}{lllrrr}
\caption{\label{kstars} Sample stars with absolute magnitude}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endfirsthead
\caption{continued.}\\
\hline\hline
Catalogue& $M_{V}$ & Spectral & Distance & Mode & Count Rate \\
\hline
\endhead
\hline
\endfoot
%%
Gl 33    & 6.37 & K2 V & 7.46 & S & 0.043170\\
Gl 66AB  & 6.26 & K2 V & 8.15 & S & 0.260478\\
Gl 68    & 5.87 & K1 V & 7.47 & P & 0.026610\\
         &      &      &      & H & 0.008686\\
Gl 86
\footnote{Source not included in the HRI catalog. See Sect.~5.4.2 for details.}
         & 5.92 & K0 V & 10.91& S & 0.058230\\
\end{longtable}
\end{landscape}
}
%
%-------------------------------------------------------------
%               Appendices have to be placed at the end, after
%                                        \end{thebibliography}
%-------------------------------------------------------------
%\end{thebibliography}
